\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-02-16 2-the-cubble-pkg.Rmd
flesh out section 2.1 - 2.4

\textcolor{red}{more beginning} In cubble data can represented in two formats: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes. The decision on which form to use in data manipulation is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier. The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of time invariant variables, or an operation that summarises time varying variables into site. Underneath the nested form, a cubble is built from a \code{rowwise_df} data frame where each site forms a separate group. This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{purrr::map()} syntax when working with list-column.

For those operations whose output involves both a spatial and temporal identifier, long form should be used. The long form uses a  \code{grouped_df} data frame to forms all the time of a site as a group. Time invariant variables are stored separately as an special attribute of the long cubble. This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.

@@

\textcolor{red}{more beginning} A cubble can be created with \code{as_cubble()} by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude. The naming of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package. \textcolor{red}{give some context of the data about to be used: a subset of a larger data etc} Below is an example of creating a cubble and by default, the cubble created is in the nested form:

@@

There are a few information in the cubble header: the name of the \code{key} variable: \code{id} and its number: 5, the bounding box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.

@@

The long cubble is suitable to manipulate the time dimension of the data and  \code{stretch()} switches the nested cubble into the long cubble. It first extracts all the spatial variables into a separate tibble and then unnests the \code{ts} column. The spatial variables are then attached as an attributes of the long cubble and can be accessed through `spatial()`:

```{r echo = TRUE}
cubble_long <- cubble_nested %>% stretch(ts)
cubble_long

spatial(cubble_long)
```

Notice that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.

@@

Analysts may need the time invariant and varying variable to be in the same table to apply certain function. In cubble, this can be done through \code{migrate()}, which moves the time invariant variables into the long cubble:

@@

Notice that these added time invariant variables will not be remembered by cubble and will disappear if switched to the nested form and then switched back. Hence, \code{migrate()} should usually be used as the last step of the analysis:

@@

\textcolor{red}{A beginning sentence that summarises the subsection}. 
\textcolor{red}{Given more details and examples when going through each subsection}

>>>>

\textcolor{red}{more beginning} In cubble data can represented in two formats: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes. The decision on which form to use in data manipulation is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier. The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of time invariant variables, or an operation that summarises time varying variables into site. Underneath the nested form, a cubble is built from a \code{rowwise_df} class where each site forms a separate group. This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{purrr::map()} syntax when working with list-column.

For those operations whose output involves both a spatial and temporal identifier, long form should be used. The long form uses a  \code{grouped_df} class to forms all the time of a site as a group. Time invariant variables are stored separately as an special attribute of the long cubble. This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.

@@

To use functions in the `cubble` package, an analyst will need to first turn the data object into a \code{cubble} class. \code{as_cubble()} does this through supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude. The use of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package. The cubble created by default is in the nested form.

Before the formal examples in Section \ref{examples}, each function introduced in this section will be demonstrated with a short example with a data that contains 5 Australia weather stations, a subset of hundreds of weather stations in Australia. This dataset contains spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information: date, maximum and minimum temperature and precipitation for 2020. \code{climate_flat} stores this data in format 1) in Figure \ref{fig:illu-input} and the code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:

@@

There are a few information in the cubble header: the name of the \code{key} variable: \code{id} and its unique number: 5, the bounding box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.

@@

The nested format is convenient for those operations whose output doesn't contain a time dimension, for those that are identified by both the spatial and temporal identifier, a long cubble is more suitable. Analysts will need to switch the nested cubble into a long cubble for such operation and \code{stretch()} is designed for this purpose: it first extracts all the spatial variables into a separate tibble and then unnests the \code{ts} column. This code switches the nested cubble just created into its long form:

```{r echo = TRUE}
cubble_long <- cubble_nested %>% stretch(ts)
cubble_long
```

Notice that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}. The spatial variables are still part of the data object but as a \code{spatial} attribute:

```{r}
attr(cubble_long, "spatial")
```

@@

Sometimes, analysts may need the spatial and temporal variable to be in the same table to apply some variable transformation. Cubble allows this operation through \code{migrate()}, which  moves the specified spatial variables into the long cubble:

@@

This operation should generally be used in the last step of the analysis since it is a temporary operation, meaning meaning these added spatial variables will disappear if switched to the nested form and then switched back:

@@

The last 4 subsections introduces operations specific to the \code{cubble} class and this section will demonstrate how the \code{cubble} class interacts with existing packages, specifically, \code{dplyr}, \code{tsibble}, \code{sf}, and \code{netcdf4}.

### 2022-02-16 main.tex
flesh out section 2.1 - 2.4

\textcolor{red}{more beginning} A cubble can be created with
\code{as_cubble()} by supplying the three key components: \code{key} as
the spatial identifier; \code{index} as the temporal identifier; and a
vector of \code{coords} in the order of longitude and latitude. The
naming of \code{key} and \code{index} follows the convention in the
\pkg{tsibble} package.
\textcolor{red}{give some context of the data about to be used: a subset of a larger data etc}
Below is an example of creating a cubble and by default, the cubble
created is in the nested form:

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested <- cubble::climate_flat %>%
+   as_cubble(key = id, index = date, coords = c(long, lat))
R> cubble_nested
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

There are a few information in the cubble header: the name of the
\code{key} variable: \code{id} and its number: 5, the bounding box:
\code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable
nested in the \code{ts} column with its type:
\code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.

@@

The long cubble is suitable to manipulate the time dimension of the data
and \code{stretch()} switches the nested cubble into the long cubble. It
first extracts all the spatial variables into a separate tibble and then
unnests the \code{ts} column. The spatial variables are then attached as
an attributes of the long cubble and can be accessed through
\texttt{spatial()}:

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_long <- cubble_nested %>% stretch(ts)
R> cubble_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\begin{CodeInput}
R> spatial(cubble_long)
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 5 x 6
  id            lat  long  elev name           wmo_id
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl>
1 ASN00009021 -31.9  116.  15.4 perth airport   94610
2 ASN00010311 -31.9  117. 179   york            94623
3 ASN00010614 -32.9  117. 338   narrogin        94627
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120
5 ASN00015131 -17.6  134. 220   elliott         94236
\end{CodeOutput}
\end{CodeChunk}

Notice that the third line in the header now shows the name and type of
spatial variables:
\code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_back <- cubble_long %>% tamp()
R> cubble_back
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

Analysts may need the time invariant and varying variable to be in the
same table to apply certain function. In cubble, this can be done
through \code{migrate()}, which moves the time invariant variables into
the long cubble:

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig <- cubble_long %>% migrate(long, lat)
R> cubble_mig
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

Notice that these added time invariant variables will not be remembered
by cubble and will disappear if switched to the nested form and then
switched back. Hence, \code{migrate()} should usually be used as the
last step of the analysis:

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig %>% tamp() %>% stretch()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # casting a tsibble into cubble
R> dt <- climate_flat %>% 
+   tsibble::as_tsibble(key = id, index = date) %>% 
+   cubble::as_cubble(coords = c(long, lat))
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # add station-based features in the nested form.
R> dt %>% 
+   mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts      tmax_mean tmax_var
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>      <dbl>    <dbl>
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_t~      25.7    38.6 
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_t~      26.2    51.1 
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_t~      23.7    45.4 
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_t~      33.1     3.02
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_t~      34.6    24.7 
\end{CodeOutput}
\end{CodeChunk}

>>>>

To use functions in the \texttt{cubble} package, an analyst will need to
first turn the data object into a \code{cubble} class.
\code{as_cubble()} does this through supplying the three key components:
\code{key} as the spatial identifier; \code{index} as the temporal
identifier; and a vector of \code{coords} in the order of longitude and
latitude. The use of \code{key} and \code{index} follows the convention
in the \pkg{tsibble} package. The cubble created by default is in the
nested form.

Before the formal examples in Section \ref{examples}, each function
introduced in this section will be demonstrated with a short example
with a data that contains 5 Australia weather stations, a subset of
hundreds of weather stations in Australia. This dataset contains spatial
information of each station: station id, latitude, longitude, elevation,
station name and World Meteorology Organisation ID and also daily
temporal information: date, maximum and minimum temperature and
precipitation for 2020. \code{climate_flat} stores this data in format
1) in Figure \ref{fig:illu-input} and the code below creates a cubble
out of \code{climate_flat} with \code{id} as the key, \code{date} as the
index, and \code{c(long, lat)} as the coordinates:

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested <- cubble::climate_flat %>%
+   as_cubble(key = id, index = date, coords = c(long, lat))
R> cubble_nested
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

There are a few information in the cubble header: the name of the
\code{key} variable: \code{id} and its unique number: 5, the bounding
box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of
variable nested in the \code{ts} column with its type:
\code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.

@@

The nested format is convenient for those operations whose output
doesn't contain a time dimension, for those that are identified by both
the spatial and temporal identifier, a long cubble is more suitable.
Analysts will need to switch the nested cubble into a long cubble for
such operation and \code{stretch()} is designed for this purpose: it
first extracts all the spatial variables into a separate tibble and then
unnests the \code{ts} column. This code switches the nested cubble just
created into its long form:

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_long <- cubble_nested %>% stretch(ts)
R> cubble_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

Notice that the third line in the header now shows the name and type of
spatial variables:
\code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}. The
spatial variables are still part of the data object but as a
\code{spatial} attribute:

\begin{CodeChunk}
\begin{CodeOutput}
# A tibble: 5 x 6
  id            lat  long  elev name           wmo_id
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl>
1 ASN00009021 -31.9  116.  15.4 perth airport   94610
2 ASN00010311 -31.9  117. 179   york            94623
3 ASN00010614 -32.9  117. 338   narrogin        94627
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120
5 ASN00015131 -17.6  134. 220   elliott         94236
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_back <- cubble_long %>% tamp()
R> cubble_back
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

Sometimes, analysts may need the spatial and temporal variable to be in
the same table to apply some variable transformation. Cubble allows this
operation through \code{migrate()}, which moves the specified spatial
variables into the long cubble:

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig <- cubble_long %>% migrate(long, lat)
R> cubble_mig
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

This operation should generally be used in the last step of the analysis
since it is a temporary operation, meaning meaning these added spatial
variables will disappear if switched to the nested form and then
switched back:

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig %>% tamp() %>% stretch()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # casting a tsibble into cubble
R> dt <- climate_flat %>% 
+   tsibble::as_tsibble(key = id, index = date) %>% 
+   cubble::as_cubble(coords = c(long, lat))
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # add station-based features in the nested form.
R> dt %>% 
+   mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts      tmax_mean tmax_var
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>      <dbl>    <dbl>
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_t~      25.7    38.6 
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_t~      26.2    51.1 
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_t~      23.7    45.4 
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_t~      33.1     3.02
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_t~      34.6    24.7 
\end{CodeOutput}
\end{CodeChunk}
